### 2024-09-23 Analysis of bias in arrests/paper/paper.qmd
Initial Commit

---
title: "No Evidence of Bias in Arrests Made by the Toronto Police Service"
author: "Aviral Bhardwaj"
abstract: The ethnic identity of citizens arrested by police has consistently sparked discussion, highlighting skewness in certain ethnicities. This paper analyzes Toronto Police Service arrest data from 2020 to 2021, considering factors such as age and search on arrest. While population distribution varies significantly, findings suggest no inherent bias in arrests. However, strict enforcement is consistent across all ethnicities and races, even when considering evidence of higher crime rates in specific groups. Further analysis can provide insights for policymakers, law enforcement, and community advocates, emphasizing the need for transparency and accountability in policing.
date: today
date-format: long
number-sections: true
format: pdf
toc: true
thanks: "Code and data supporting this analysis is available at: https://github.com/Aviral-03/Ethnic-Identity-Arrests-Toronto"
bibliography: references.bib
---

```{r}
#| echo: false
# All the required packages
## Installing Packages only needs to be done once

#install.packages("tidyverse")
#install.packages("tinytex")
#install.packages("opendatatoronto")
#install.packages("kableExtra")
#install.packages("patchwork")
#install.packages("janitor")
#install.packages("gridExtra")
```

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(opendatatoronto)
library(tinytex)
library(knitr)
library(dplyr)
library(ggplot2)
library(janitor)
library(patchwork)
library(kableExtra)
library(gridExtra)
```

```{r}
#| warning: false
#| message: false
#| include: false

# Get package data

# get package
package <- show_package("police-race-and-identity-based-data-collection-arrests-strip-searches")
package

# get all resources for this package
resources <- list_package_resources("police-race-and-identity-based-data-collection-arrests-strip-searches")

# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))

# load the first datastore resource as a sample
data <- filter(datastore_resources, row_number()==1) %>% get_resource()

# Save the data in CSV
write.csv(data, "../../inputs/data/raw_data.csv")
```

## Introduction

In the vibrant metropolis of Toronto, the intersection between law enforcement practices and demographic characteristics holds profound implications for community dynamics and public policy. Toronto, named Canada's most diverse city, encounters numerous challenges arising from the coexistence of ethnically diverse populations within the same neighborhoods. According to an article published by Georgetown University, the "ongoing debate surrounding police reforms is heavily influenced by historical connections with slavery and contemporary racial disparities" (@rJackson2022). This statement underscores the contentious nature of whether the race identity affiliation of police officers affects their policing practices, despite the expectation for fair and unbiased treatment.

In a recent article published by the Toronto Police Services, it is highlighted that "Black, East/Southeast Asian, Middle Eastern, and Latino people were over-represented in reported use of force incidents compared to their presence in the enforcement action population" (@Fanfair2022). This indicates that, despite Toronto Police Services working towards a more equitable society, there are still hints of racial bias in the arrests made by the Toronto Police Services.Conversely this paper undertakes a comprehensive analysis of arrest data of various race and ethinic iden spanning the years 2020 and 2021 in the city of Toronto, seeking to unravel the intricate patterns that shape the city's legal landscape.

Arrest data serves as a vital lens through which we can gain insights into the dynamics of law enforcement activities and their correlation with ethnicity and age. In the data section we clean and process our data. Following which in our discussion and results our inquiry begins with a broad exploration of the distribution of arrests across various age groups, aiming to discern overarching trends and fluctuations. The age spectrum, a fundamental dimension of this analysis, reveals intriguing patterns that underscore the nuanced nature of arrests in Toronto. This paper aims to contribute to understanding the dynamics of arrests in Toronto, providing insights that can inform targeted interventions, policy considerations, and community engagement efforts.

## Data

This data is sourced from the Toronto Open Data Portal `opendatatoronto` [@rOpenDataToronto]. The following data is used for an in-depth analysis of arrests and strip searches which involve various ethnic and identity-based groups in the city of Toronto. The information is gathered by the Toronto Police Service (TPS) and is collected under the authority of the Police Services Act. Its purpose is to improve the relationship between the police and the community by being more transparent with the public. The data covers the period from 2020 to 2021.

The data underwent cleaning and analysis using the R programming language [@r]. Cleaning was performed with the `tidyverse` package [@rTidyverse], involving the removal of unnecessary columns for the analysis. Subsequently, analysis was conducted utilizing the `dplyr` package [@rDplyr], with the data grouped by ethnic and identity-based categories. Visualization was then executed using the `ggplot2` package [@rGgplot2], depicting the number of arrests and strip searches across various ethnic and identity-based groups.

The raw dataset comprised 32,000 arrest records spanning both 2020 and 2021, containing 26 variables of analyzable data. However, for the analysis in this paper, we narrowed our focus to specific variables, including Arrest Year, Perceived Race, Age Group, Youth at Arrest, Booked, Search Reason Cause Injury, Search Reason Assist Escape, Search Reason Possess Weapons, and Search Reason Possess Evidence. Additionally, we consolidated similar columns, namely Search Reason Cause Injury, Search Reason Assist Escape, Search Reason Possess Weapons, and Search Reason Possess Evidence, into a unified column labeled "Search Reason." utilizing the `any()` function [@rDplyr].

```{r}
#| warning: false
#| message: false
#| include: false

# Reading the data
data <- read.csv("../../inputs/data/raw_data.csv")

# Adjusting and cleaning the name of the variable
cleaned_data <- data %>% 
  clean_names() %>% 
  clean_names(case = "snake")

cleaned_data <- cleaned_data |>
  select(
    arrest_year, perceived_race, age_group_at_arrest, youth_at_arrest_under_18_years, booked, search_reason_cause_injury, search_reason_assist_escape, search_reason_possess_weapons, search_reason_possess_evidence)

# Combining the columns that are similar
cleaned_data <- cleaned_data %>%
  mutate(search_reason = case_when(
    search_reason_cause_injury == 1 ~ "1",
    search_reason_assist_escape == 1 ~ "1",
    search_reason_possess_weapons == 1 ~ "1",
    search_reason_possess_evidence == 1 ~ "1",
    TRUE ~ "0"
  ))

# Rename the columns
cleaned_data <- cleaned_data %>%
  rename(
    age_group = age_group_at_arrest,
    youth_at_arrest_under_18_years = youth_at_arrest_under_18_years
  )

# Mutating the date for age groups involving Aged 17 years and younger and Aged 17 years and under essentially they mean the same thing
cleaned_data <- cleaned_data %>%
  mutate(age_group = case_when(
    age_group == "Aged 17 years and younger" ~ "Aged 17 years and under",
    age_group == "Aged 17 years and under" ~ "Aged 17 years and under",
    TRUE ~ age_group
  ))

# Mutating for booked with NA values
cleaned_data <- cleaned_data %>%
  mutate(booked = case_when(
    is.na(booked) ~ "0",
    booked == 1 ~ "1",
    TRUE ~ "Not Booked"
  ))

cleaned_data <- cleaned_data |>
  select(
    arrest_year, perceived_race, age_group, youth_at_arrest_under_18_years, booked, search_reason)

write.csv(cleaned_data, "../data/cleaned_data.csv")

```

```{r}
#| warning: false
#| message: false
#| include: false
# Make a table of the data with the number of arrests and strip searches by ethinic and identity based groups for the year 2020, and 2021
# cleaned_data <- read.csv("../data/cleaned_data.csv")

# List of the ethnic and identity based groups that are in the data
# ethnic_backgrounds <- cleaned_data |> select(perceived_race) |> unique()

# List of the years that are in the data
# years <- cleaned_data |> select(arrest_year) |> unique()

# List of the age groups that are in the data
# age_groups <- cleaned_data |> select(age_group) |> unique()

# write.csv(cleaned_data, "../data/cleaned_data.csv")
```

### Arrests by Ethnic and Identity-Based Groups for Year 2020 and 2021

Since our analysis mainly focuses on the distribution of arrests for each ethnic identity, we extracted all ethnic identities from our data. These ethnic identities include Black, White, South Asian, East Asian, Middle Eastern, Latin American, Southeast Asian, and Indigenous. The data for the years 2020 and 2021 are summarized in Table 1 (a) (@tbl-table1) and Table 1 (b), respectively.

Table 1 (a) (

@tbl-table1) provides a summary of our data, revealing the number of individuals arrested from each major ethnic group in the years 2020 and 2021. Table 2 (b) shows, out of all the people who were arrested from each ethnic identity, how many of them were actually booked under a criminal case.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-table1
#| tbl-cap: "Total number of arrests vs Booked for the year 2020, and 2021"
## Table 1: Total Number of Arrests
cleaned_data <- read.csv("../data/cleaned_data.csv")

table1 <- cleaned_data |> 
  filter(!perceived_race == "None") |>
  group_by(arrest_year, perceived_race) |> 
  summarise(n = n(), .groups = "drop") |> 
  arrange(desc(n)) |> 
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

table2 <- cleaned_data |>
  filter(booked == 1) |>
  filter(!perceived_race == "None") |>
  group_by(arrest_year, perceived_race) |>
  summarise(n = n(), .groups = "drop") |>
  arrange(desc(n)) |>
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

knitr::kable(
  list(table1, table2),
  col.names = c("Ethinic Identity", "2020", "2021"),
  booktabs = TRUE,
  )

```

### Total number of arrests by age group for the year 2020 and 2021

It would also be important to take a look at the age group distribution of arrests. Table 2 (@tbl-table2) shows us this information. A data summary which indicates the count of different age groups of individuals from prominent ethnic groups in the years 2020 and 2021. All the age groups were the following: Aged 17 years and under, Aged 18 to 24 years, Aged 25 to 34 years, Aged 35 to 44 years, Aged 45 to 54 years, Aged 55 to 64 years, Aged 65 years and over.

Due to the extensive nature of the dataset, going through nine different age groups across eight ethnicities, summarizing the information for clarity is extremely difficult. Consequently, we have decided to showcase a subset of the data which focuses on the age group of 25 to 34 years so that it is easy to read.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-table2
#| tbl-cap: "Arrests Booked by ethinic background for the year 2020, and 2021"

## Table 3: Arrests by Age Group

cleaned_data <- read.csv("../data/cleaned_data.csv")
# 
# # Group 
table3 <- cleaned_data %>%
  filter(!(age_group == "None") & !(perceived_race == "None")) |>
  group_by(arrest_year, perceived_race, age_group) %>%
  summarise(n = n(), .groups = "drop") %>%
  arrange(desc(perceived_race)) %>%
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

table3_small <- cleaned_data %>%
  filter(age_group == "Aged 25 to 34 years") |>
  group_by(arrest_year, perceived_race, age_group) %>%
  summarise(n = n(), .groups = "drop") |>
  arrange(desc(n)) |>
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

kable(table3_small,
      align = "c",
      linesep = "",
      col.names = c("Ethinic and Identity Based Groups", "Age Group", "2020", "2021"),
      booktabs = TRUE)

```

We would like to acknowledge that opendatatoronto [@rOpenDataToronto] provides the distribution of the population by ethnicity. However, due to the vast size of Toronto and its numerous wards, it has been difficult to compare all of the data and cross-verify it with our crime rate data comprehensively. Despite this, a more detailed analysis can be conducted if we compare specific ethnic groups with each other.

Now that the data has been cleaned, we can begin to analyze the data. The following section will provide a summary of the data, and the trends that emerge from the data. Figure 1 (@fig-fig1), Figure 2 (@fig-fig2), Figure 3 (@fig-fig3), and Figure 4 (@fig-fig4) was created using `ggplot` [@rGgplot2], displays the information that can be used as compare the trends as described above.

### Number of Arrests in year 2020-2021

With reference to Table 1 (@tbl-table1) and the graph (see @fig-fig1) the data shows that there is a non-uniform distribution of arrests across all ethnic groups, with the Black ethnic group and White group having the highest numbers of arrests in both 2020 and 2021. In addition to this, the data shows a decrease in the number of arrests for these two ethnic groups from 2020 to 2021. However, a  for the other ethnic groups, the number of arrests has increased over the same time period. This shows that both the trends are moving in opposite directions.

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| fig.cap: "Number of arrests by ethinic background for the year 2020, and 2021"
#| fig.width: 10
#| fig.height: 6
#| fig.align: 'center'
#| fig.pos: 'htb'
#| label: fig-fig1

table1_long <- pivot_longer(table1, cols = c("2020", "2021"), names_to = "year", values_to = "value")

# Create a plot
ggplot(table1_long, aes(x = perceived_race, y = value, fill = year)) +
  geom_bar(stat = "identity", position = 'dodge') +
  labs(title = "Bar Chart for the number of arrests by ethinic background for the year 2020, and 2021",
       x = "Ethnic and Identity Based Groups",
       y = "Number of Arrests",
       fill = "Year") +
  theme_minimal() +
  scale_fill_discrete(labels = c("2020", "2021")) +
  theme(legend.position = "right") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_brewer(palette = "Set1")
```

Referring to Table 1 (@tbl-table1) for the proportion of arrests that were booked, we plotted a stacked bar graph (see @fig-fig2), with y-axis representing the total percentage. The data shows that there is a consistent distribution of the number of individuals who were booked after arrests across various ethnic groups. Notably, there are many more individuals who were booked after being arrested as  compared to those who were arrested but not booked for a criminal case. In addition to this, we also observe that, across all ethnic groups, the percentage of individuals booked falls within the range of 50% to 55%, and this pattern remains consistent in the year 2021.

However the Black ethnicity has the highest number of individuals who were booked after being arrested in both the years 2020 and 2021.

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| fig.cap: "Number of arrests booked by ethinic background for the year 2020, and 2021"
#| fig.width: 10
#| fig.height: 6
#| fig.align: 'center'
#| fig.pos: 'htb'
#| label: fig-fig2

table2_long <-
  cleaned_data |>
  select(
    arrest_year,
    perceived_race,
    booked
  ) |>
  group_by(arrest_year, perceived_race, booked) %>%
  summarise(n = n(), .groups = "drop") %>%
  arrange(desc(perceived_race)) %>%
  filter(!(perceived_race == "None")) %>%
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

table2_long <- pivot_longer(table2_long, cols = c("2020", "2021"), names_to = "year", values_to = "value")

ggplot(table2_long, aes(x = perceived_race, y = value, fill = factor(booked))) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~ year, scales = "free_y") +
  labs(title = "Booked after Arrest in the year 2020 and 2021",
       x = "Perceived Race",
       y = "Percentage of Arrests Booked",
       fill = "Status") +
  theme_minimal() +
  theme(legend.position = "right") + 
  scale_fill_discrete(labels = c("Not Booked", "Booked")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Number of Arrests by Age Group for the year 2020-2021

As depicted in Figure 3 (see @fig-fig3), the data shows there is a non-uniform distribution of arrests across age groups, with the age groups of 25 to 34 years and 35 to 44 years showing the highest number of arrests in both 2020 and 2021. Furthermore, there is a notable decrease in the number of arrests for these two age groups from 2020 to 2021. However, a different trend emerges for other age groups, where the number of arrests has increased over the same time period.

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| fig.cap: "Number of arrests by Age-Group for the year 2020, and 2021"
#| fig.width: 10
#| fig.height: 6
#| fig.align: 'center'
#| fig.pos: 'htb'
#| label: fig-fig3
# Bar plot

table3_long <- pivot_longer(table3, cols = c("2020", "2021"), names_to = "year", values_to = "value")

ggplot(table3_long, aes(x = perceived_race, y = value, fill = factor(age_group))) +
  geom_bar(stat = "identity") +
  facet_wrap(~ year, scales = "free_y") +
  labs(title = "Bar Chart for the number of arrests by Age-Group for the year 2020, and 2021",
       x = "Perceived Race",
       y = "Number of People",
       fill = "Year") +
  theme_minimal() +
  theme(legend.position = "bottom") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Examining Figure 4 (see @fig-fig4), which shows the dot plot for the number of arrests by age group in the year 2020 and 2021 for the top two ethnic groups, "Black" and "White," it becomes evident that the age group 25 to 34 years represents the highest number of arrests for individuals of Black ethnicity. The trend for Black ethnicity indicates a decreasing curve from this age group onward. Howver, for individuals of White ethnicity, the peak in arrests is observed in the age group 35 to 44 years. From age 45 to 65, however, the difference in the number of arrests between both ethnicities starts to converge.

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| fig.cap: "Scatter Plot for the number of arrests by Age-Group for the year 2020, and 2021"
#| fig.width: 10
#| fig.height: 6
#| fig.align: 'center'
#| fig.pos: 'htb'
#| label: fig-fig4

table3_ii <- table3 %>%
  filter(perceived_race %in% c("White", "Black")) %>%
  filter(!(age_group == "None")) %>%
  pivot_longer(cols = c("2020", "2021"), names_to = "year", values_to = "value")

# Create a line Graph
ggplot(table3_ii, aes(x = age_group, y = value, color = perceived_race)) +
  geom_point(size = 3) +
  facet_wrap(~ year, scales = "free_y") +
  labs(title = "Scatter Plot for the number of arrests by Age-Group for the year 2021",
       x = "Age Group",
       y = "Number of People",
       color = "Perceived Race") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Discussion

The analysis presented in this paper aimed to shed light on the potential bias in arrests made by the Toronto Police Service, specifically focusing on ethnic and identity-based groups. The findings from the data analysis provide valuable insights into the dynamics of law enforcement activities in the city.

### Lack of Inherent Bias in Arrests

Our analysis suggests no inherent bias in arrests made by the Toronto Police Service. The distribution of arrests across various ethnic and identity-based groups does not show a consistent pattern of discrimination. This observation is crucial, as it challenges the narrative surrounding racial bias in policing and suggests that strict enforcement is consistent across all ethnicities.

### Age Group Dynamics

The examination of age group dynamics in arrests provides additional nuances to the analysis. The age groups of 25 to 34 years and 35 to 44 years emerge as having the highest number of arrests in both 2020 and 2021. Interestingly, there is a noticeable decrease in arrests for these age groups from 2020 to 2021, suggesting potential shifts in law enforcement patterns.

The detailed exploration of age groups within specific ethnicities, such as the comparison between "Black" and "White" ethnicities, reveals intriguing patterns. The peak in arrests for individuals of Black ethnicity occurs in the age group of 25 to 34 years, while for individuals of White ethnicity, the peak is observed in the age group of 35 to 44 years. This divergence in age-specific arrest patterns highlights the importance of considering intersectionality in understanding policing dynamics.

### Implications for Policymaking and Community Engagement

Even though this paper provides several essential insights, we must acknowledge that it has its limitations. In order to understand the data better, a more detailed comparison of specific ethnic groups and further research is required. Law enforcement agencies and community advocates can gain from these insights by planning intervention targeted at specific parts of the population and community engagement efforts. This would also allow the policymakers to make any policy adjustments that they may feel is necessary to improve the situation. The emphasis on transparency and accountability in policing remains crucial, as highlighted in report by the Toronto Police Services Board [@rTPSB2019].

###Conclusion
In conclusion, tha analysis of the arrest data in the city of Toronto shows that despite there being concerns about racial bias. The data suggests that arrests may be driven by objective factors rather than racial profiling. The reforms brought about by the Toronto Police Service have strengthened transparency and fairness within the police force.While disparities may exist due to socio-economic factors, the data does not show any systemic bias.

## References